---
title: "General Social Survey"
author: "mariahlynn"
date: "Year 2024"
output: 
  html_document:
    number_sections: true
    toc: true
---


```{r include=FALSE,echo=FALSE}
require(tigerstats)
require(tidyverse)
require(foreign)
source("GrabData.R")
```




#  Question 1

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

The data in the dataframe GSSdata is from the 2018 General Social Survey. The first blocks of R-code has selected down a subset of the data to just 16 variables. It has further removed unwanted factor levels in much of the data. Examine the code in the GrabData.R file to see what it is doing. Some of the variables are categorical and others are numerical. Be sure to do a variable analysis before tackling each question.  

First question - Is opinion on the death penalty (CAPPUN) independent of gun ownership (OWNGUN)?

###Introduction      

$H_0$: CAPPUN is independent of OWNGUN

$H_A$: They are dependent 

## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

CAPPUN is a xxxxxxxx variable, (categorical variable)
OwNGUN is a yyyyyyy variabl, so I
use the zzzzzz helper file

### Results  

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

#### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->


```{r}
barchartGC(~CAPPUN + OWNGUN,data=GSSdata)
barchartGC(~CAPPUN + OWNGUN,data=GSSdata, type="percent")

```

This graph shows that half of the people who own guns are in favor of them, while more than half of the people who do not own guns are not in favor of them.   
  

#### Numerical Descriptive Results

```{r}
table1<- xtabs(~CAPPUN + OWNGUN,data=GSSdata)
rowPerc(table1)
colPerc(table1)
```

This table shows that the relationship between gun ownership(OWNGUN) and opinions on capital punishment(CAPPUN), indicate that a higher percentage of gun owners favor capital punishment compared to non-gun owners.

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->



### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The p-value for the Chi-squared test was p= 0.01651
This means we reject/fail to reject the null hypothesis and conclude that opinion on capital punishment (CAPPUN) is not independent of gun ownership(OWNGUN) 


 the p-value or for the fisher exact test was p=xxxx, and the odds ratio was yyyyy
 so the odds ratio,2.271:1 means that peope who do not own a gun are 2,271 times more likely to oppose capital punishment 


#  Question 2

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

Is there a difference in years of education (EDUC) between the those who did or did not vote in 2012 (VOTE12)?


###Introduction     
$H_0$:There is a difference in years of educations between those who did not vote in 2012.

$H_A$: There is no difference between education if someone voted. 

## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

EDUC is categorical, VOTE12 is also categorical. 

We want EDUC ~ VOTE12
The helper file we use is CAT~CAT

###Results  

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

#### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
barchartGC(~CAPPUN + OWNGUN,data=GSSdata)
barchartGC(~CAPPUN + OWNGUN,data=GSSdata,type="percent")
```
This shows that there was an increase of people who voted in 2012 tended to have higher education 

#### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
table1<-xtabs(~CAPPUN + OWNGUN,data=GSSdata)
rowPerc(table1)
colPerc(table1)
```

The numerical study shows that those who favor CAPPUN are 75% gun owners. 

### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
t.test(EDUC~VOTE12, data=GSSdata)
```
This T-test shows that those who have voted had a better education. Meaning we have to reject the null hypothesis.
The P-value 0.000001248 is less than 0.5. Meaning we reject the null hypothesis. 

#  Question 3

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

Determine if a persons performance on the wordsum test (WORDSUM) is independent of their level of schooling (DEGREE).

###Introduction 
$H_0$: Performance on the wordsum test is independent of their level of schooling  

$H_A$: Performance on the wordsum test is a dependent of their level of schooling  


## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

WORDSUM is a numerical variable,
DEGREE is a categorical variable, so I use the num~cat data file. 

##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->



### Descriptive Results

#### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
boxplot(WORDSUM~DEGREE,data=GSSdata)
```

The boxplot shows that as the degree level increases from LT High School to Graduate, the median WORDSUM scores also increase. Graduates tend to have higher WORDSUM scores compared to those with lower degrees.  

#### Numerical Descriptive Results


```{r}
favstats(WORDSUM~DEGREE, data=GSSdata) 
```

The numerical result shows that as the level of education increases.The WORDSUM scores tend to increase. Graduates have the highest median and mean WORDSUM scores, followed by those with Bachelor degrees, Junior College degrees, High School degrees, and finally those with less than High School education.

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->


### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
anova=aov(WORDSUM~DEGREE,data=GSSdata)
summary(anova)
```
We can see that the P-value is 0.00000025 which is smaller than 0.05. Therefore, we have to reject the null hypothesis

```{r}
#we specify the margins starting at the bottom and proceeding clockwise-thus 8.7 is the left side margin to hold the names of the factors.
par(mar = c(7, 8.7, 2, 2)) 
plot(TukeyHSD(anova) ,las=2)
```

There is no difference between high school and less than high school.  
There is a difference between junior college and high school.   
There is a difference between bachelor college and high school.    
There is a difference between graduate and high school.     
There is no difference between highschool and graduate.     
There is no difference between junior college and graduate.    
There is no difference between bachelor and graduate.      

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variables causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->


#### Question 1: Is gun ownership independant of opinion on captal punishment. 

The summary results and the inferential results show that they are not independent.

#### Question 2: Is there a difference in years of education (EDUC) between the those who did or did not vote in 2012 (VOTE12)?

The summary results and the inferential results show that there is a difference of years in people who voted in 2012. 

#### Question 3: Is a persons performance on the wordsum test (WORDSUM) independent of their level of schooling (DEGREE)?

The summary results and the inferential results show that a person's performance on the WORDSUM test (WORDSUM) is not independent of their level of schooling (DEGREE).


